---
title: "Competition_Model"
author: "JT_Miller"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Proposed question: Can we model competition between specialist and generalist bees on pollen resource?

#### Things I want to incorporate in the model

      1) Generalist bees are polylectic, and therefore are able to utilize multiple flower types for pollen extraction
      2) Specialist bees are monolectic, and therefore will only be able to utilize a singular flower type for pollen extraction
      3) Generalist will be more "clumsy" at pollen extraction, meaning that per time step, they extract less pollen then their specialist competitors. 
      

```{r}
# Load in some libraries
library(tidyverse)
```

```{r}
# Define the parameters
phi <- 100 # The conversion of flowers to pollen 
Ug <- 0.05 # The uptake rate of pollen for generalist bees
Us <- 0.1 # The uptake rate of pollen for specialist bees
hg <- 1 # The handling time of pollen for generalist (how long to take it back to brood)
hs <- 1 # The handling time of pollen for specialist (how long to take it back to brood)
mg <- 0.7  # The mortality of the generalist bees
ms <- 0.7  # The mortality of the specialist bees
rmax_e <- 0.05 # The maximum reproductive rate potential of the Exotic plant 
rmax_d <- 0.05 # The maximum reproductive rate potential of the Native plant
delta <- 2 # A proportion of YS that makes plants based upon pollen yield
Ke <- 100 # The carrying capacity of the exotic plants 
Kd <- 100 # The carrying capacity of the native plants 
alpha <- 0.05 # competition coefficient of D’s effect on E
beta <- 0.05  # competition coefficient of E’s effect on D
x <- 100 # required number of pollen grains to have max reproduction. 
Yg <- 0.005   # The yield conversion of generalist bees to pollen
Ys <- 0.01  # The yield conversion of specialist bees to pollen

# Define the initial conditions 
G0 <- 1 # Initial population of generalist bees
S0 <- 1 # Initial population of specialist bees
E0 <- 1 # Initial population of Exotic plants
D0 <- 1 # Initial population of Native plants
A0 <- 10 # Initial amount of alien pollen
N0 <- 10 # Initial amount of native pollen 
```

```{r}
### The Model 

tset <- seq(from = 0, to = 2500, length.out = 10000) # Create a vector of time steps

# Create a holding vector and fill the initial conditions 
G.simu <- NaN*tset; G.simu[1] <- G0
S.simu <- NaN*tset; S.simu[1] <- S0
E.simu <- NaN*tset; E.simu[1] <- E0
D.simu <- NaN*tset; D.simu[1] <- D0
A.simu <- NaN*tset; A.simu[1] <- A0
N.simu <- NaN*tset; N.simu[1] <- N0

# For each time step
for(i in 2:length(tset)){
  
  # Calc change in time
  dt <- tset[i]-tset[i-1]
  
  # store placeholder variables
	N <- N.simu[i-1]
	A <- A.simu[i-1]
	G <- G.simu[i-1]
	S <- S.simu[i-1]
	E <- E.simu[i-1]
	D <- D.simu[i-1]
	
	# Calc the pollens
	dN <- (D*phi - (1/Yg)*((Ug*N*G)/(1 + Ug*hg*N + Ug*hg*A)) - (1/Ys)*((Us*N*S)/(1+Us*hs*N)))*dt
	
	dA <- (E*phi - (1/Yg)*((Ug*A*G)/(1 + Ug*hg*N + Ug*hg*A)))*dt
	
	# Calc the bees
	dG <- ((G*(Ug*A + Ug*N)/(1 + Ug*hg*N + Ug*hg*A)) - mg*G)*dt
	dS <- (((Us*N*S)/(1+Us*hs*N)) - ms*S)*dt
	
	# Calc the plants
	
	dE <- (rmax_e*(min(1, G/x)) * max(0,((Ke - E - alpha*D)/Ke)))*dt
	
	dD <- (rmax_d*(min(1, (delta*S + G)/x))*max(0,((Kd - D - beta*E)/Kd)))*dt
	
	### E > 0 ###
	### D > 0 ###
	
	# Calc totals 
	N.simu[i] <- N + dN
	A.simu[i] <- A + dA
	G.simu[i] <- G + dG
	S.simu[i] <- S + dS
	E.simu[i] <- E + dE
	D.simu[i] <- D + dD
	
} # End of for loop

# Plot it out
plot(x = tset, y = N.simu,
     type = 'l', las = 1, lwd = 1, col = "green",
     xlab = "time", ylab="Amount",
     ylim = c(-100000, 100000))
lines(x = tset, y = A.simu,
      lwd = 2, col = "yellow")
# Plot bees
#lines(x = tset, y = G.simu,
      #lwd = 2, col = "blue")
#lines(x = tset, y = S.simu,
      #lwd = 2, col = "purple")
# Plot plants
lines(x = tset, y = D.simu,
      lwd = 2, col = "green")
lines(x = tset, y = E.simu,
      lwd = 2, col = "orange")
### Current issue is that the implicit competition portion of the equation is driving the values to negative. 
```

#### Taking Holly's advice and figuring out how to model it with only the native plant and specialist pollinator involved to find where things go crazy

```{r}
# Lets think about the parameters needed just here
rmax_d <- 0.05 # The maximum reproductive rate potential of the Native plant
delta <- 2 # A proportion of YS that makes plants based upon pollen yield
x <- 100 # 10000 # required number of pollen grains to have max reproduction.
Kd <- 10 # 1000 # The carrying capacity of the native plants
Us <- .1 # The uptake rate of pollen for specialist bees
hs <- 0 # The handling time of pollen for specialist (how long to take it back to brood)
ms <- 0.2  # The mortality of the specialist bees
phi <- 10 # 10 # The conversion of flowers to pollen 
Ys <- .01# 100  # The yield conversion of specialist bees to pollen

# Initial Conditions
D0 <- 1 # Initial population of Native plants
N0 <- 10 # Intial amount of native pollen in the system
S0 <- 1


# Make the time scale
tset2 <- seq(from = 0, to = 20000, by = .01) # Change this

# Set intial values for the simulation. 
D.simu2 <- NaN*tset2; D.simu2[1] <- D0
N.simu2 <- NaN*tset2; N.simu2[1] <- N0
S.simu2 <- NaN*tset2; S.simu2[1] <- S0

# For each time step
for(i in 2:length(tset2)){
  
  # Calc change in time
  dt <- tset2[i]-tset2[i-1]
  
  # store placeholder variables
	D <- D.simu2[i-1]
  N <- N.simu2[i-1]
	S <- S.simu2[i-1]
	
	# Calc the number of native plants 
  dD <- ((rmax_d*(min(1, (S*delta)/x)))*D*(1-(D/Kd)))*dt
  
  dN <- (D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N)))*dt
    #if(N <= 0){dN <- 0} else {dN <- dN}
    #if(N <= 0){N <- 0} else {N <- N}
  dS <- (((Us*N*S)/(1+Us*hs*N)) - ms*S)*dt
  
  # Calc totals 
  D.simu2[i] <- D + dD
	N.simu2[i] <- N + dN
	S.simu2[i] <- S + dS
	
	
} # End of for loop

plot(x = tset2, y = N.simu2,
     type = 'l', las = 1, lwd = 1, col = "orange",
     xlab = "time", ylab="Relative abundance") 


plot(x = tset2, y = D.simu2,
     type = 'l', las = 1, lwd = 1, col = "orange",
     xlab = "time", ylab="Relative abundance") 


plot(x = tset2, y = S.simu2,
     type = 'l', las = 1, lwd = 1, col = "orange",
     xlab = "time", ylab="Relative abundance") 
     #ylim = c(0, 1000))
#lines(x = tset2, y = D.simu2,
     # lwd = 2, col = "green")
#lines(x = tset2, y = S.simu2,
      #lwd = 2, col = "purple")


######################## All on one graph #######################
plot(x = tset2, y = S.simu2,
     type = 'l', las = 1, lwd = 1, col = "orange",
     xlab = "time", ylab="Relative abundance",
     ylim=c(0, max(D.simu2)))
lines(x = tset2, y = D.simu2,
      lwd = 2, col = "green")
lines(x = tset2, y = N.simu2,
      lwd = 2, col = "purple")

```
### Check whats happening with some bifurcation diagrams

#### We know that mortality seems to be a big driver in pollen equilibrium, so lets start there. 
```{r}
# Set a range of mortality for the specialist bee
ms_set <- seq(from=0.01, to = 0.99, by = 0.01)

# Set up some empty values for our pollen equilibrium for the same length as ms_set
Nstarset <- rep(NaN, times = length(ms_set))
Dstarset <- rep(NaN, times = length(ms_set))
Sstarset <- rep(NaN, times = length(ms_set))

# Set up a time interval
timepoints <- seq(from = 1, to = 100, by = 0.01)

# Nest a for-loop
for(j in 1:length(ms_set)){
  
  m <- ms_set[j]
  
  N.simu <- rep(NaN, times = length(timepoints))
  D.simu <- rep(NaN, times = length(timepoints))
  S.simu <- rep(NaN, times = length(timepoints))
  
  N.simu[1] <- N0
  D.simu[1] <- D0
  S.simu[1] <- S0
  
  for(i in 2:length(timepoints)){ 
    # store previous population size as N
    N <- N.simu[i-1]
    D <- D.simu[i-1]
    S <- S.simu[i-1]
    
    # calculate change in time
    deltat <- timepoints[i] - timepoints[i-1] 

    # calculating the change in population size
    deltaD <- ((rmax_d*(min(1, (S*delta)/x)))*D*(1-(D/Kd))) * deltat 
    deltaN <- (D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N))) * deltat
    deltaS <- (((Us*N*S)/(1+Us*hs*N)) - m*S) * deltat 
    
    # calculating total population size
    D.simu[i] = D + deltaD
    N.simu[i] = N + deltaN
    S.simu[i] = S + deltaS
  }
  Nstarset[j] <- N.simu[length(timepoints)] 
  Dstarset[j] <- D.simu[length(timepoints)]
  Sstarset[j] <- S.simu[length(timepoints)]
}


```

#### And plot it 
```{r}
plot(x = ms_set, y = Nstarset, 
     type='l', 
     lwd=2, 
     xlab = "Mortality Rate of Specialist Bees", 
     ylab = "Stable Equilibria of Pollen, N*")

plot(x = ms_set, y = Dstarset, 
     type='l', 
     lwd=2, 
     xlab = "Mortality Rate of Specialist Bees", 
     ylab = "Stable Equilibria of Plants, D*")

plot(x = ms_set, y = Sstarset, 
     type='l', 
     lwd=2, 
     xlab = "Mortality Rate of Specialist Bees", 
     ylab = "Stable Equilibria of Specialist Bees, S*")
```










### Creating a function to run our model
```{r}
Specialist.Only <- function(D0, S0, N0, phi, Ys, hs, ms, rmax_d, Kd){
  tset2 <- seq(1, 100, length.out=1000)
  
  D.simu2 <- NaN*tset2; D.simu2[1] <- D0
  N.simu2 <- NaN*tset2; N.simu2[1] <- N0
  S.simu2 <- NaN*tset2; S.simu2[1] <- S0
  
  for(i in 2:length(tset2)){
  
  # Calc change in time
  dt <- tset2[i]-tset2[i-1]
  
  # store placeholder variables
	D <- D.simu2[i-1]
  N <- N.simu2[i-1]
	S <- S.simu2[i-1]
	
	# Calc the number of native plants 
  dD <- ((rmax_d*(min(1, (S*delta)/x)))*D*(1-(D/Kd)))*dt
  
  dN <- (D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N)))*dt
    
  dS <- (((Us*N*S)/(1+Us*hs*N)) - ms*S)*dt
  
  # Calc totals 
  D.simu2[i] <- D + dD
	N.simu2[i] <- N + dN
	S.simu2[i] <- S + dS
	
	
  }
  out<-list(tset2=tset2,D.simu2=D.simu2, N.simu2=N.simu2, S.simu2=S.simu2 )
  return(out)
  
}


simu3 <- Specialist.Only(1,1,10,100,0.01,1,0.7,0.05,100)

d_data <- simu3$D.simu2
n_data <- simu3$N.simu2
s_data <- simu3$S.simu2

plot(d_data,type='l',col="red",xlab="Time",ylab="Population Size")
lines(n_data,col='blue')#plot predator
#legend(1,3,legend=c("V","P"),col=c("red","blue"),lty=c(1,1))
```

### Using multiroot() to solve for the equilibria
```{r}
library(rootSolve)
Simplified_Bees = function(x, parms){
  D = x[1]
  N = x[2]
  S = x[3]
  rmax_d = parms[1]
  Kd = parms[2]
  Us = parms[3]
  hs = parms[4]
  Ys = parms[5]
  phi = parms[6]
  ms = parms[7]
  
  plants = rmax_d*D*(1-(D/Kd))
  pollen = D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N))
  sp.bees = ((Us*N*S)/(1+Us*hs*N)) - ms*S
  
  return(c(plants, pollen, sp.bees))
}

# Specify the parameters 
parameters <- c(rmax_d = 0.05, Kd = 10, Us = 0.1, hs = 0, Ys = 0.01, phi = 10, ms = 0.2)

# Guesses for our three state variables
x = c(D = 1, N = 10, S = 1)

# Call in multiroot
eqm = multiroot(Simplified_Bees, x, parms = parameters)

# Look at the roots
eqm$root
```

