---
title: "Competition_Model"
author: "JT_Miller"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Proposed question: Can we model competition between specialist and generalist bees on pollen resource?

#### Things I want to incorporate in the model

      1) Generalist bees are polylectic, and therefore are able to utilize multiple flower types for pollen extraction
      2) Specialist bees are monolectic, and therefore will only be able to utilize a singular flower type for pollen extraction
      3) Generalist will be more "clumsy" at pollen extraction, meaning that per time step, they extract less pollen then their specialist competitors. 
      

```{r}
# Load in some libraries
library(tidyverse)
```
```{r}
# Define the parameters
phi <- # The conversion of pollen to bees
Ug <-  # The uptake rate of pollen for generalist bees
Us <-  # The uptake rate of pollen for specialist bees
hg <-  # The handling time of pollen for generalist (how long to take it back to brood)
hs <-  # The handling time of pollen for specialist (how long to take it back to brood)
mg <-   # The mortality of the generalist bees
ms <- # The mortality of the specialist bees
rmax_e <- # The maximum reproductive rate potential of the Exotic plant 
rmax_d <- # The maximum reproductive rate potential of the Native plant
delta <- # A proportion of YS that makes plants based upon pollen yield
Ke <- # The carrying capacity of the exotic plants 
Kd <- # The carrying capacity of the native plants 
alpha <- # competition coefficient of D’s effect on E
beta <- # competition coefficient of E’s effect on D
x <- # required number of pollen grains to have max reproduction. 
  
# Define the initial conditions 
G0 <- 100 # Initial population of generalist bees
S0 <- 100 # Initial population of specialist bees
E0 <- 10000 # Initial population of Exotic plants
D0 <- 10000 # Initial population of Native plants
```

```{r}
### The Model 

tset <- seq(from = 0, to = 10000, length.out = 10000) # Create a vector of time steps

# Create a holding vector and fill the initial conditions 
G.simu <- NaN*tset; G.simu[1] <- G0
S.simu <- NaN*tset; S.simu[1] <- S0
E.simu <- NaN*tset; E.simu[1] <- E0
D.simu <- NaN*tset; D.simu[1] <- D0


```

      