---
title: "Model_ODEs"
author: "JT Miller"
date: "5/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### This Markdown is built to solve the ODEs for the bee model using tools I've learned in EEMB 247 Quantitative Biology

```{r}
# Necessary Libraries
library(deSolve)
library(rootSolve)
```


### Here I want to describe the system looking just at how a specialist bee interacts with native flowers in pollen uptake.
#### Assumptions 
    i. The bee is uptakes pollen at a steady rate
    ii. The pollen uptaken by the bee benefits the plant by contributing to its maximum potential for reproduction (rmax)
    iii. The amount of plants present in the system determines how much pollen is being produced. 
***
# N is the pollen
# S is the specialist bees
# D is the native plants
$dNdt = D*phi - (1/Ys)((Us N  S/ (1+Us hs N)))$
$dSdt = ((Us N S)/(1+Us hs N)) - ms N)$
$dDdt = rmax (min(1, (s*delta)/x)) D (1-(D/Kd))$

***
### Analytical Solutions for these equations at equilibrium
$D^* = Kd$
$N^* = ms/(1 - hs*ms)Us$
$S^* = (D(1+hs*N*Us)Ys*phi)/N*Us$

### Solving for the equilibria Numerically
```{r}
# Model
Bee.model <- function(t,x,params){
  # State Variables
  N = x[1]
  S = x[2]
  D = x[3]
  # Parameters
  phi = params[1]
  Ys = params[2]
  Us = params[3]
  hs = params[4]
  ms = params[5]
  rmax = params[6]
  Kd = params[7]
  xreq = params[8]
  delta = params[9]
  
  # Set up the equations
  dNdt = D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N))
  dSdt = ((Us*N*S)/(1+Us*hs*N)) - ms*S
  dDdt = rmax*((S*delta/xreq)/(1 + (S*delta/xreq)))*D*(1-(D/Kd))
  
  # Return the output
  return(list(c(dNdt, dSdt, dDdt)))
  
  
}

# Set up the parameters
phi = 100 # The amount of pollen produced by each plant
Ys = 0.01 # The yield of pollen for each bee
Us = 0.01 # The uptake of the pollen by each bee
hs = 1 # The handling time for bee transporting the pollen
ms = 0.7 # The mortality of the bees per time step
rmax = 0.05 # The maximum potential for reproduction of the plants
Kd = 100 # The Carrying Capacity of the native plants
xreq = 100 # The number of required pollenation visits to meet maximum reproductive potential
delta = 2 # Specialist bees are worth 2x the amount of a generalist bee that Im using as baseline for these equations

# Set up initial values for the state variables, lets also try doing this with equilibrium values to see if analytical effect was correct
Dstar = Kd
Nstar = ms/(Us*(1-ms*hs))
Sstar = (Dstar*Ys*phi)/ms

D0 = Dstar
N0 = Nstar
S0 = Sstar



# Set up the time
times <- seq(0, 1000, by = 0.01) 

# Initial Values
initial_values = c(N0,S0,D0)



# Parameters
Parameters = c(phi, Ys, Us, hs, ms, rmax, Kd, xreq, delta)

Bee.model(times, initial_values, Parameters)

results = lsoda(initial_values, times, Bee.model, Parameters)

colnames(results) = c("time", "N", "S","D")

results


# Plotting in base plot
plot(results[, "time"], results[, "N"], type="l", col="orange",  
     xlab="time", ylab="Quantity", ylim=c(0, max(results[, "N"]) + 100)) 

lines(results[, "time"], results[, "S"], type="l", col="purple")  

lines(results[, "time"], results[, "D"], type="l", col="green")  

legend("topright", legend=c("N", "S", "D"), fill=c("orange", "purple", "green"))

```

# Simulation to check how long it takes to reach the analytical equilibriums
### Solving for the equilibria Numerically
```{r}
# Model
Bee.model <- function(t,x,params){
  # State Variables
  N = x[1]
  S = x[2]
  D = x[3]
  # Parameters
  phi = params[1]
  Ys = params[2]
  Us = params[3]
  hs = params[4]
  ms = params[5]
  rmax = params[6]
  Kd = params[7]
  xreq = params[8]
  delta = params[9]
  
  # Set up the equations
  dNdt = D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N))
  dSdt = ((Us*N*S)/(1+Us*hs*N)) - ms*S
  dDdt = rmax*((S*delta)/xreq)*D*(1-(D/Kd))
  
  # Return the output
  return(list(c(dNdt, dSdt, dDdt)))
  
  
}

# Set up the parameters
phi = 100 # The amount of pollen produced by each plant
Ys = 0.01 # The yield of pollen for each bee
Us = 0.01 # The uptake of the pollen by each bee
hs = 1 # The handling time for bee transporting the pollen
ms = 0.7 # The mortality of the bees per time step
rmax = 0.05 # The maximum potential for reproduction of the plants
Kd = 100 # The Carrying Capacity of the native plants
xreq = 1000 # The number of required pollenation visits to meet maximum reproductive potential
delta = 2 # Specialist bees are worth 2x the amount of a generalist bee that Im using as baseline for these equations


D0 = 150
N0 = 235
S0 = 160



# Set up the time
times <- seq(0, 100, by = 1) 

# Initial Values
initial_values = c(N0,S0,D0)



# Parameters
Parameters = c(phi, Ys, Us, hs, ms, rmax, Kd, xreq, delta)

Bee.model(times, initial_values, Parameters)

results = lsoda(initial_values, times, Bee.model, Parameters)

colnames(results) = c("time", "N", "S","D")

results


# Plotting in base plot
plot(results[, "time"], results[, "N"], type="l", col="orange",  
     xlab="time", ylab="Quantity", ylim=c(0, max(results[, "S"]) + 100)) 

lines(results[, "time"], results[, "S"], type="l", col="purple")  

lines(results[, "time"], results[, "D"], type="l", col="green")  

legend("topright", legend=c("N", "S", "D"), fill=c("orange", "purple", "green"))

```

```{r}
# The function used by multiroot to calculate
Bee.model.eqms <- function(x, parms){
  # State Variables
  N = x[1]
  S = x[2]
  D = x[3]
  # Parameters
  phi = parms[1]
  Ys = parms[2]
  Us = parms[3]
  hs = parms[4]
  ms = parms[5]
  rmax = parms[6]
  Kd = parms[7]
  xreq = parms[8]
  delta = parms[9]
  
  c(F1 = D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N)),
    F2 = ((Us*N*S)/(1+Us*hs*N)) - ms*S,
    F3 = rmax*(Min(1,(S*delta)/xreq))*D*(1-(D/Kd)))
  
}

# Function returning the gradients of the model (used in Jacobian.full)

Bee.model.full <- function(t,x,parameters){
  # State Variables
  N = x[1]
  S = x[2]
  D = x[3]
  # Parameters
  phi = parameters[1]
  Ys = parameters[2]
  Us = parameters[3]
  hs = parameters[4]
  ms = parameters[5]
  rmax = parameters[6]
  Kd = parameters[7]
  xreq = parameters[8]
  delta = parameters[9]
  
  # Set up the equations
  dNdt = D*phi - (1/Ys)*((Us*N*S)/(1+Us*hs*N))
  dSdt = ((Us*N*S)/(1+Us*hs*N)) - ms*S
  dDdt = rmax*(min(1,(S*delta)/xreq))*D*(1-(D/Kd))
  
  # Return the output
  return(list(c(dNdt, dSdt, dDdt)))
  
  
  
  
}
# Initial Values
x <- c(N0 = 233.33,S0 = 142.857,D0 = 100)
# Parameters
parameters <-  c(phi, Ys, Us, hs, ms, rmax, Kd, xreq, delta)

#call multiroot
eqm = multiroot(Bee.model.eqms, x, parms = parameters)

#the equilibrium values are stored in the root part of eqm
eqm$root

# Numerically Calculate the Jacobian matrix at the equilibrium. 
JAC = jacobian.full(y=c(eqm$root), func = Bee.model.full, parms = parameters)
JAC
```


